Theorems and Proofs

### 2017-05-28 Thresholding Bandit/Theorem1.tex
Improve TBP Theorem 1

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\newtheorem{theorem}{Theorem}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}

@@

In Theorem 1 of Locatelli et al. (2016), a lower bound for the
thresholding bandit problem is shown. The basic idea is as follows. The
problem formulation considered in the paper describes bandit problems
for Sub-Gaussian distributions. To show a lower bound, we show that
every algorithm is expected to make a certain error even if we restrict
ourselves to a heavily specified scenario. Even then, any algorithm will
have a certain level of risk with which it cannot distinguish between
the actual setting and a slightly changed setting (under which the
former answer leads to regret).

To be more specific, consider a case in which we set \(\tau = 0\) and
\(\epsilon = 0\) w.l.o.g. If we then choose all arms to have normally
distributed feedback with means \(\mu_i\), we see that for each arm
\(i\) the distance from the threshold is given by
\(\Delta_i = |\mu_i - \tau| + \epsilon = \mu_i\). Thus, assuming a
variance of 1 for every arm, we have for every arm the distribution
\(\nu_i = \mathcal{N}(\mu_i,1) = \mathcal{N}(\Delta_i,1)\). Furthermore,
define \(\nu_i' = \mathcal{N}(-\Delta_i,1)\).

The idea now is that every arm \(i \in \{1,\dots,K\}\) generates a
sample for every time point \(t\in \{1, \dots, T\}\) from its
distribution \(\nu_i\) before the algorithm starts to choose arms. Thus,
we have a \(T \times K\) table of realized observations from the product
distribution given by
\(\mathcal{B}^0 = \nu_1^0 \otimes \ldots \otimes \nu_K^0 = \nu_1^0 \otimes \ldots \otimes \nu_K\).
This is a product distribution where the \(K\) different means all lie
above the threshold. We would hope that an algorithm classifies the
means accordingly. However, we will show in what follows that any
algorithm makes a certain error by not being able to distinguish
\(\mathcal{B}^0\) from a product distribution \(\mathcal{B}^i\) which
differs only slightly and has the same overall complexity (as defined by
\(H = \sum_{i=1}^K (\Delta_i)^{-2}\)).

Before the start of the algorithm, we introduce a slight variation in
the setup. Similar to how one can switch arms in the pure exploration
setup for multi-armed bandits (compare Audibert et al., Garivier \&
Kaufmann) to ``confuse'' the algorithm without introducing additional
complexity to the specified problem. However, since we do not compare
arms with each other directly in the thresholding bandit problem
(switching arms would still keep all means above the threshold), we
instead need to flip an arm with respect to the threshold. Since the
threshold is \(\tau = 0\), we have for the flipped arm the new
distribution \(\nu_i'\) as defined above. Indeed, we can define the new
product distribution \(B^i = \nu_1^i \otimes \dots \otimes \nu_K^i\)
where for
\(k \leq K, \nu_k' := \nu_i \mathbbm{1}_{k \neq i} + \nu_i' \mathbbm{1}_{k=i}\).
This flipping of the arm corresponds to multiplying all random variables
\(X_{i,t}\) of arm \(i\) by \(-1\).

@@

Given our formulation of the regret, the expected regret
\(\mathbb{E}[\mathcal{L}(T)]\) corresponds directly to the probability
of making a wrong classification of any of the \(K\) means. In general,
let \(T\) be a binary decision function, and define
\(\Omega = \{T = 0\}\) and \(\Omega^C = \{T = 1\}\). Assume that the
decision \(T\) distinguishes between two states of the world, \(H_0\)
and \(H_1\). Then our overall error probability is given by
\(P_{H_0}(T=1) + P_{H_1}(T = 0) = P_{H_0}(\Omega^C) + P_{H_1}(\Omega)\).

In our case, \(H_0\) may correspond to the state in which we have
\(\mathcal{B}^0\), and \(H_1\) corresponds to the case where any arm has
been flipped, \(\mathcal{B}^i\). Furthermore, we consider the events in
which the algorithm, making the decision, classifies arm \(i\) as being
above the threshold: \(\mathcal{A}_i = \{i \in \hat{S}_\tau\). The event
on which all arms are being classified as above the threshold is
consequently given by \(\mathcal{A} = \bigcap \mathcal{A}_i\). However,
Theorem 1 does not bound the overall regret, but the maximum error
probability. We can write this now as

\begin{equation}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}^C) \big)
= \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \label{LocatelliTheorem1ExpRegret}
\end{equation}

In order to give a bound on above equation, we start by considering
\(\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)\). This probability can be
measured as the expected change in the likelihood of observed data times
the probability under the original state. The measure of event
\(\mathcal{A}_i\) under state \(\mathcal{B}^i\) is given as the
probability of \(\mathcal{A}_i\) under the (original) state
\(\mathcal{B}^0\) multiplied by a factor. This factor is the change in
the data likelihoods due to moving from \(\mathcal{B}^0\) to
\(\mathcal{B}^i\), and is expressed as the expected likelihood ratio
under \(\mathcal{B}^0\). By definition though, this is equal to the
empirical Kullback-Leibler divergence of the two measures.

First, let us define the change of measure in terms of the empirical
log-likelihood for our case. Then, we will show how to replace the
log-likelihood by the empirical Kullback-Leibler divergence.

@@

\paragraph{Change of Measure}\label{change-of-measure}

We are now ready to state the change of measure. As alluded to above, we
consider how the probability changes for the event \(\mathcal{A}_i\)
when we move from \(\mathcal{B}^0\) to \(\mathcal{B}^i\). When flipping
arm \(i\), we only changed the samples of this arm, that is, the first
\(T_i\) observations we got from arm \(i\) the algorithm took. The
change of measure is then defined as:

@@

\paragraph{Kullback-Leibler
Divergence}\label{kullback-leibler-divergence}

For our problem, the Kullback-Leibler divergence of \(/nu_k'\) from
\(\nu_k\) is given by \(\KL_k := KL(\nu_k', \nu_k) = 2\Delta_k^2\).

@@

@@

itself is a normally distributed random variable with mean zero and a
variance decreasing in \(t\). Thus, one can show that
\(\mathbb{P}_{\mathcal{B}^i}(\xi) \geq 3/4\) for the event

\begin{equation}
\xi = \{ \forall k \leq K, \forall t \leq T, |\hat{\KL}_{k,t} - \KL_{k,t}| \leq 4 \Delta_k \sqrt{\frac{\log(4(\log(T)+1)K)}{t}} \}. \label{LocatelliTheorem1EventXi}
\end{equation}

@@

Part of our bound is
\(\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)\),
so let's continue with that. In general it holds that
\(\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \bar{X}\). Thus we can
write:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & \geq \frac{1}{K} \sum_{i=1}{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) \\
& \geq \frac{1}{K} \sum_{i=1}{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i \cap \xi) \\
& \geq \frac{1}{K} \sum_{i=1}{K} \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 2 \Delta_i^2 T_i - 4 \Delta_i \sqrt{T_i} \sqrt{\log(4(\log(T)+1)K)}) ]
\end{align*}

where the second line is again the intersection of \(\mathcal{A}_i\) and
\(\xi\) as before in order to apply the change of measure again in the
third line.

@@

\paragraph{Conclusion}\label{conclusion}

Consider again the risk which we need to bound from equation
\eqref{LocatelliTheorem1ExpRegret}. We bound the maximum expected loss
over all \(\mathcal{B}_i\). By again using the fact that in general
\(\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \bar{X}\), we plug in
our previous calculations:

>>>>

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}

@@

In Theorem 1 of Locatelli et al. (2016), a lower bound for the
thresholding bandit problem is shown. The basic idea is as follows. The
problem formulation considered in the paper describes bandit problems
for sub-Gaussian distributions. To show a lower bound, we show that
every algorithm is expected to make a certain error even if we restrict
ourselves to a very restricted scenario. Even then, any algorithm will
have a certain level of risk with which it cannot distinguish between
the actual setting and slightly changed settings (under which the former
answer leads to regret).

To be more specific, consider a case in which we set \(\tau = 0\) and
\(\epsilon = 0\) w.l.o.g. If we then choose all arms to have normally
distributed feedback with means \(\mu_i\), we see that for each arm
\(i\) the distance from the threshold is given by
\(\Delta_i = |\mu_i - \tau| + \epsilon = \mu_i\). Thus, assuming a
variance of 1 for every arm, we have for every arm the distribution
\(\nu_i = \mathcal{N}(\mu_i,1) = \mathcal{N}(\Delta_i,1)\). Furthermore,
define for each arm \(i \in \{1, \dots, K\}\) the alternative
distribution \(\nu_i' = \mathcal{N}(-\Delta_i,1)\). It is easy to see
that in this case, this is equivalent to a ``flip around the
threshold''. Arms that originally have a mean larger than the threshold
have a mean smaller than the threshold after the flip.

The idea now is that every arm \(i \in \{1,\dots,K\}\) generates a
sample for every time point \(t\in \{1, \dots, T\}\) from its
distribution \(\nu_i\) before the algorithm starts to choose arms. Thus,
we have a \(T \times K\) table of realized observations from the product
distribution given by
\(\mathcal{B}^0 = \nu_1^0 \otimes \ldots \otimes \nu_K^0 = \nu_1^0 \otimes \ldots \otimes \nu_K\).
This is a product distribution where the \(K\) different means all lie
above the threshold. We would hope that an algorithm classifies the
means accordingly. However, we will show in what follows that any
algorithm makes a certain error by not being able to distinguish
\(\mathcal{B}^0\) from at least one of \(K\) alternative product
distributions \(\mathcal{B}^i, i \in \{1, \dots, K\}\), of which each
differs only slightly in one arm from the original distribution, and has
the same overall complexity (as defined by
\(H = \sum_{i=1}^K (\Delta_i)^{-2}\)). We require \(K\) different
distributions so that an arm sampling arm \(i\) very often for some
reason will not have a smaller lower bound by chance.

For each of the \(K\) arms, we introduce an alternative model of the
following kind. Before the start of the algorithm, we introduce a slight
variation in the setup. The idea is similar to how one can switch arms
in the pure exploration setup for multi-armed bandits (compare Audibert
et al., 2010; Garivier \& Kaufmann, 2016) to ``confuse'' any algorithm
without introducing additional complexity to the specified problem.
However, since we do not compare arms directly with each other in the
thresholding bandit problem (switching arms would still keep all means
above the threshold), we instead need to flip an arm \(i\) with respect
to the threshold in order to get the alternative model
\(\mathcal{B}_i\). Since the threshold is \(\tau = 0\), we have for the
flipped arm the new distribution \(\nu_i'\) as defined above. Indeed, we
can define the new product distribution
\(B^i = \nu_1^i \otimes \dots \otimes \nu_K^i\) where for
\(k \leq K, \nu_k' := \nu_i \mathbbm{1}_{k \neq i} + \nu_i' \mathbbm{1}_{k=i}\).
This flipping of the arm corresponds to multiplying all random variables
\(X_{i,t}\) of arm \(i\) by \(-1\).

@@

Given our formulation of the regret, the expected regret
\(\mathbb{E}[\mathcal{L}(T)]\) corresponds directly to the probability
of making a wrong classification of any of the \(K\) means. In general,
let \(g\) be a binary decision function, and define
\(\Omega = \{g = 0\}\) and \(\Omega^C = \{g = 1\}\). Assume that the
decision \(T\) distinguishes between two states of the world, \(H_0\)
and \(H_1\). Then our overall error probability is given by
\(P_{H_0}(g=1) + P_{H_1}(g = 0) = P_{H_0}(\Omega^C) + P_{H_1}(\Omega)\).

In our case, \(H_0\) may correspond to the state in which we have
\(\mathcal{B}^0\), and \(H_1\) corresponds to the case where any arm has
been flipped, \(\mathcal{B}^i\). Furthermore, we consider the events in
which the algorithm, making the decision, classifies arm \(i\) as being
above the threshold: \(\mathcal{A}_i = \{i \in \hat{S}_\tau\}\). The
event on which all arms are being classified as above the threshold is
consequently given by \(\mathcal{A} = \bigcap \mathcal{A}_i\). However,
Theorem 1 does not bound the overall regret, but the maximum error
probability on the \(K+1\) different models. We can write this now as

\begin{align}
\max_{i \in \{0, \dots, K\}} \mathbb{E}_{\mathcal{B}^i} (\mathcal{L}(T)) & \geq \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}^C) \big) \\
& = \max \big( \max_{i \in \{1, \dots, K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i), 1 - \mathbb{P}_{\mathcal{B}^0}(\mathcal{A}) \big) \label{LocatelliTheorem1ExpRegret}
\end{align}

In order to give a bound on above equation, we start by considering
\(\mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)\). This probability can be
written as the expected change in the likelihood of observed data when
moving from distribution \(P_{\mathcal{B}_0}\) to \(P_{\mathcal{B}_i}\),
times the probability under the original state. In other words, the
probability of event \(\mathcal{A}_i\) under state \(\mathcal{B}^i\) is
given as the probability of \(\mathcal{A}_i\) under the (original) state
\(\mathcal{B}^0\) multiplied by a factor. This factor describes whether
the new distribution increases or decreases the probability of the
event, and is given by the change in the data likelihoods due to moving
from \(\mathcal{B}^0\) to \(\mathcal{B}^i\), expressed as the expected
likelihood ratio under \(\mathcal{B}^0\). By definition, the latter is
equal to the empirical Kullback-Leibler divergence of the two
distributions.

We will define the change of distribution in terms of the empirical
log-likelihood for our case. Then, we will show how to replace the
log-likelihood by the empirical Kullback-Leibler divergence. So let us
first introduce the empirical Kullback-Leibler divergence.

@@

\paragraph{Change of Distribution}\label{change-of-distribution}

We are now ready to state the change of distribution. As alluded to
above, we consider how the probability changes for the event
\(\mathcal{A}_i\) when we move from \(\mathcal{B}^0\) to
\(\mathcal{B}^i\). When flipping arm \(i\), we only changed the samples
of this arm, that is, the first \(T_i\) observations we got from arm
\(i\) the algorithm took. The change of distribution is then defined as:

@@

@@

For our problem, the Kullback-Leibler divergence of \(\nu_k'\) from
\(\nu_k\) is given by \(\KL_k := KL(\nu_k', \nu_k) = 2\Delta_k^2\).

@@

itself is a normally distributed random variable with mean zero and a
variance decreasing in \(t\). Thus, one can show as in Locatelli et al.
(2016) that \(\mathbb{P}_{\mathcal{B}^i}(\xi) \geq 3/4\) for the event

\begin{equation}
\xi = \{ \forall k \leq K, \forall t \leq T, |\hat{\KL}_{k,t} - \KL_{k,t}| \leq 4 \Delta_k \sqrt{\frac{\log(4(\log(T)+1)K)}{t}}\}. \label{LocatelliTheorem1EventXi}
\end{equation}

@@

Part of our bound is
\(\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i)\),
so let's continue with that. In general it holds that
\(\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \frac{1}{K}\sum_{i=1}^K X_i\).
Thus we can write:

\begin{align*}
\max_{i \in \{1,\dots,K\}} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) & \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i) \\
& \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{P}_{\mathcal{B}^i}(\mathcal{A}_i \cap \xi) \\
& \geq \frac{1}{K} \sum_{i=1}^{K} \mathbb{E}_{\mathcal{B}^0} \big[\mathbbm{1}_{\mathcal{A}_i \cap \xi} \exp (- 2 \Delta_i^2 T_i - 4 \Delta_i \sqrt{T_i} \sqrt{\log(4(\log(T)+1)K)}) ]
\end{align*}

where the second line is again the intersection of \(\mathcal{A}_i\) and
\(\xi\) as before in order to apply the change of distribution in the
third line.

@@

\paragraph{Bringing everything
together}\label{bringing-everything-together}

Consider again the risk which we need to bound from equation
\eqref{LocatelliTheorem1ExpRegret}. We bound the maximum expected loss
over all \(\mathcal{B}_i\). By again using the fact that in general
\(\max_{1 \leq i \leq K} (X_1, \dots, X_K) \geq \frac{1}{K} \sum_{i = 1}^K X_i\),
we plug in our previous calculations:
